### 2021-11-08 maps/SpatialJoin/spatial_join.Rmd
Initial commit

---
title: "Spatial Join"
output: html_notebook
---

Showing how spatial join is done i sf.

```{r setup}
library(tidyverse)
library(sf)
library(here)
```

```{r}
kc_houses <- here("maps/SpatialJoin/kc_house_data_251264.gpkg") %>% 
  st_read() %>% 
  st_transform(2926)
```

```{r}
names(kc_houses)
```

```{r}
kc_tracts10 <- here("maps/SpatialJoin/tracts10.shp") %>% 
  st_read() %>% 
  st_transform(2926)
```

```{r}
names(kc_tracts10)
```

```{r}
kc_wadoh_map <- here("maps/SpatialJoin/kc_wadoh_map.gpkg") %>% 
  st_read() %>% 
  st_transform(2926)
```

```{r}
names(kc_wadoh_map)
```

```{r}
kc_tracts10_env_data <- left_join(
  kc_tracts10,
  st_drop_geometry(kc_wadoh_map),
  by = "GEO_ID_TRT"
  )
```

```{r}
names(kc_tracts10_env_data)
```

```{r}
kc_houses_env_var <- st_join(kc_houses, kc_tracts10_env_data)
```

```{r}
names(kc_houses_env_var)
```

```{r}
st_write(kc_houses_env_var, "kc_houses_env_var.gpkg", append = FALSE)
```